Toward \earlywarning\ detection of gravitational waves from compact binary coalescence

### 2011-07-11 introduction.tex
made spacing between numbers and units consistent
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1283 ccdd05d5-56c1-461a-bf98-85f6876523b4

Compact binary coalescence (\CBC{}) is a plausible progenitor for most short
gamma-ray bursts (short \GRB{}s) \citep{Lee:2005, nakar07}, but the
association is not iron-clad \citep{2011ApJ...727..109V}. The tidally
disrupted material falls onto the newly formed, rapidly spinning compact object
and is accelerated in jets along the spin axis with a timescale of
$0.1$--$1$~s after the merger \citep{Janka1999}, matching the short \GRB{}
duration distribution well. Prompt \EM{} emission including the \GRB{}
can arise as fast outflowing
matter collides with slower matter ejected earlier in inner shocks. The same
inner shocks, or potentially reverse shocks, can produce an accompanying
optical flash \citep{Sari99}. The prompt emission is a probe into the extreme
initial conditions of the outflow, in contrast with afterglows, which arise in
the external shock with the local medium and are relatively insensitive to
initial conditions. Optical flashes have been observed for a handful of long
\GRB{}s \citep{2011CRPhy..12..255A} by telescopes with extremely rapid response
or, in the case of GRB 080319b, by pure serendipity, where several
telescopes were already observing the afterglow of another \GRB{} in the same
field of view \citep{2008Natur.455..183R}. The observed optical flashes peaked
within tens of seconds and decayed quickly. For short \GRB{} energy balance and
plasma density,
however, the reverse shock model predicts a peak flux in radio, approximately
20 minutes after the \GRB{}, but also a relatively faint optical flash
\citep{nakar07}; for a once-per-year Advanced LIGO event at $130$\,Mpc, the
radio flux will peak around $9$\,GHz at $\sim$$5$\,mJy, with emission in the
$R$-band at $\sim$$19^\mathrm{th}$ magnitude. Interestingly, roughly a quarter to half of observed short
\GRB{}s also exhibit extended X-ray emission of $30$--$100$\,s in duration
beginning $\sim$$10$\,s after the \GRB{} and carrying comparable fluence to the
initial outburst. This can be explained if the merger results in the formation
of a proto-magnetar that interacts with ejecta \citep{Bucciantini2011}. Rapid
\GW{} alerts would enable joint \EM{} and \GW{} observations to confirm
the short \GRB{}-\CBC{} link and allow the early \EM{}
observation of exceptionally nearby and thus bright events.

>>>>

Compact binary coalescence (\CBC{}) is a plausible progenitor for most short
gamma-ray bursts (short \GRB{}s) \citep{Lee:2005, nakar07}, but the
association is not iron-clad \citep{2011ApJ...727..109V}. The tidally
disrupted material falls onto the newly formed, rapidly spinning compact object
and is accelerated in jets along the spin axis with a timescale of
$0.1$--$1$~s after the merger \citep{Janka1999}, matching the short \GRB{}
duration distribution well. Prompt \EM{} emission including the \GRB{}
can arise as fast outflowing
matter collides with slower matter ejected earlier in inner shocks. The same
inner shocks, or potentially reverse shocks, can produce an accompanying
optical flash \citep{Sari99}. The prompt emission is a probe into the extreme
initial conditions of the outflow, in contrast with afterglows, which arise in
the external shock with the local medium and are relatively insensitive to
initial conditions. Optical flashes have been observed for a handful of long
\GRB{}s \citep{2011CRPhy..12..255A} by telescopes with extremely rapid response
or, in the case of GRB 080319b, by pure serendipity, where several
telescopes were already observing the afterglow of another \GRB{} in the same
field of view \citep{2008Natur.455..183R}. The observed optical flashes peaked
within tens of seconds and decayed quickly. For short \GRB{} energy balance and
plasma density,
however, the reverse shock model predicts a peak flux in radio, approximately
20 minutes after the \GRB{}, but also a relatively faint optical flash
\citep{nakar07}; for a once-per-year Advanced LIGO event at $130$~Mpc, the
radio flux will peak around 9~GHz at $\sim$$5$~mJy, with emission in the
$R$-band at $\sim$$19^\mathrm{th}$ magnitude. Interestingly, roughly a quarter to half of observed short
\GRB{}s also exhibit extended X-ray emission of $30$--$100$~s in duration
beginning $\sim$$10$~s after the \GRB{} and carrying comparable fluence to the
initial outburst. This can be explained if the merger results in the formation
of a proto-magnetar that interacts with ejecta \citep{Bucciantini2011}. Rapid
\GW{} alerts would enable joint \EM{} and \GW{} observations to confirm
the short \GRB{}-\CBC{} link and allow the early \EM{}
observation of exceptionally nearby and thus bright events.
